Supplementary Figures relative to normalization

### 2020-05-12 supplementary_figures/supp2_filtering.Rmd
updated supp

```{r filt_more, fig.width=8, fig.height=9}
res <- readRDS("../data/results_filtering.rds")
m <- old_evalPlot_filtering(res, returnTable=TRUE)
m <- m[grep("veryStringent", m$method, invert=TRUE),]

# # horizontal version
# m <- m[grep("veryStringent", m$method, invert=TRUE),]
# m.f1 <- reshape2::dcast(m, method~dataset, value.var="F1atK")
# n <- sapply(strsplit(gsub("=",".",m.f1[,1]),"\\."),FUN=function(x) rev(x)[1])
# n[grep("=doublet", m.f1[,1])] <- paste0("doubletRemoval+",n[grep("=doublet", m.f1[,1])])
# m.pc <- reshape2::dcast(m, method~dataset, value.var="maxPCout")[,-1]
# m.f1m <- reshape2::dcast(m, method~dataset, value.var="meanF1")[,-1]
# m.pcm <- reshape2::dcast(m, method~dataset, value.var="medianPCout")[,-1]
# m.f1 <- m.f1[,-1]
# row.names(m.f1m) <- row.names(m.pcm) <- row.names(m.f1) <- row.names(m.pc) <- n
# 
# chm(list("mean F1"=m.f1m, "F1 at true \n# clusters"=m.f1, "max % lost per\nsubpopulation"=m.pc, "median % lost per\nsubpopulation"=m.pcm), scale = "column", colors = viridis::inferno(3), value_format = c("%.2f","%.2f","%.0f","%.0f"), scaleFn=list(scale,scale, scale2,scale2), cluster_columns = FALSE,  row_sortFn = function(x, na.rm){ sum(x[1:18],na.rm=TRUE)-median(x[19:36], na.rm=TRUE)/100})

mycast <- function(x, val, form=dataset~method){ 
  x <- reshape2::dcast(x, form, value.var=val)
  row.names(x) <- x[,1]
  as.matrix(x[,-1])
}
m.f1 <- mycast(m, "F1atK")
n <- sapply(strsplit(gsub("=",".",colnames(m.f1)),"\\."),FUN=function(x) rev(x)[1])
n[grep("=doublet", colnames(m.f1))] <- paste0("doubletRemoval+",n[grep("=doublet", colnames(m.f1))])
m.pc <- mycast(m, "maxPCout")
colnames(m.f1) <- colnames(m.pc) <- n
o <- order(colMeans(t(scale(t(m.f1))), na.rm=TRUE)*100-colMeans(t(scale(t(m.pc))), na.rm=TRUE))
m.f1 <- m.f1[,o]
m.pc <- m.pc[,o]
m.f1m <- mycast(m, "meanF1")[,o]
m.pcm <- mycast(m, "medianPCout")[,o]
colnames(m.f1m) <- colnames(m.pcm) <- colnames(m.f1)

names(nn) <- nn <- c("m.f1m","m.f1","m.pc","m.pcm")
cellfns <- lapply(nn, FUN=function(mname){
  if(grepl("f1",mname)){
    sp <- paste0('sprintf("%.1f", 100*',mname,'[i, j])')
  }else{
    sp <- paste0('sprintf("%.1f", ',mname,'[i, j])')
  }
  eval(parse(text=paste0('function(j, i, x, y, width, height, fill){',
      'lab <- ', sp, '; grid.text(lab, x, y, gp = gpar(fontsize = 10)) }')))
})

names(nn) <- c("mean F1", "F1 at true \n# clusters", "max % lost per\nsubpopulation", "median % lost per\nsubpopulation")
hs <- lapply(seq_along(nn), FUN=function(i){
  mat <- get(nn[[i]])
  if(grepl("f1",nn[[i]])){
    mat <- t(scale(t(mat)))
  }else{
    mat <- t(scale(t(sqrt(mat))))
  }
  Heatmap(mat, row_title=names(nn)[[i]], col=viridis::inferno(100), cell_fun = cellfns[[i]], show_heatmap_legend = FALSE, cluster_columns = FALSE, cluster_rows = FALSE)
})

hs[[1]] %v% hs[[2]] %v% hs[[3]] %v% hs[[4]]
```

### Supplementary Figure `r getFigNb()`
Mean clustering F1 score per subpopulation, mean F1 at true number of clusters, as well as maximum and median proportion of excluded cells per subpopulation across various filtering strategies. Doublet removal generally improves clustering accuracy with very mild exclusion rates, even in datasets that do not have heterotypic doublets. Stringent distribution-based filtering creates large cell type biases.

@@

```{r}
res <- readRDS("../data/selection_old.rds")
res <- res[grep("deviance|seurat.variance.standardized",res$sel),]
res$filt <- as.factor(res$filt)
levels(res$filt) <- c("coding only", "no mitochondrial", "no ribosomal", "all")
res2 <- res[which(abs(res$nbClusters-truek[res$dataset])<=1),]
ll <- lapply(split(res2, res2$norm), FUN=function(x){
  m2 <- cast2(x, formula=filt~dataset, fun.aggregate=median, value.var="ARI")
})
chm(ll, scale="column", scaleTitle="z-score", sameScale=TRUE, value_format="%.2f", row_sortFn=function(x,...){ median(x,...) }, cluster_columns=FALSE)
```

@@

Impact of restricting the type of features used on the ARI of the clustering.
\newpage

>>>>

```{r filt_more, fig.width=8.5, fig.height=5}
res <- readRDS("../data/results_filtering.rds")
m <- pipeComp:::.mergeFilterOut(res$evaluation[1:2])
m <- m[grep("veryStringent",m$filt,invert=TRUE),]
res$evaluation$clustering <- res$evaluation$clustering[grep("veryStringent",res$evaluation$clustering$filt,invert=TRUE),]
ag <- aggregate(m$pc.lost, m[,c("dataset","doubletmethod","filt")], FUN=max)
colnames(ag)[4] <- "maxLost"
ag$medianLost <- aggregate(m$pc.lost, m[,c("dataset","doubletmethod","filt")], FUN=median)$x

renameHrows(evalHeatmap(ag, what=c("maxLost","medianLost"), agg.by=c("doubletmethod","filt"), 
            scale=TRUE, row_split=doubletmethod, reorder_rows=TRUE, value_format="%.f",
            title = c("max % lost per\nsubpopulation","median % lost per\nsubpopulation"), 
            font_factor = 0.75, show_column_names = TRUE, anno_legend = FALSE) + 
  evalHeatmap(res, step="clustering", what=c("mean_F1"), agg.by=c("doubletmethod","filt"), title="mean F1", 
              font_factor = 0.75, show_column_names=TRUE, anno_legend=FALSE) + 
  evalHeatmap(res, step="clustering", what=c("mean_F1"), agg.by=c("doubletmethod","filt"), 
              filterExpr = n_clus==true.nbClusts, font_factor = 0.75, show_column_names=TRUE, 
              anno_legend=FALSE, name="F1atK", title="mean F1\nat true K"),
  function(x){
    x <- gsub("^doublet\\.","",x)
  }
)
```

### Supplementary Figure `r getFigNb()`
Mean clustering F1 score per subpopulation, mean F1 at true number of clusters, as well as maximum and median proportion of excluded cells per subpopulation across various filtering strategies. Doublet removal generally improves clustering accuracy with relatively mild increases exclusion rates, even in datasets that do not have heterotypic doublets. Stringent distribution-based filtering creates large cell type biases.

@@

```{r, fig.width=8, fig.height=6.5}
res <- readRDS("../data/results_featexcl.rds")
rnfn <- function(x){
  x <- gsub(" > norm.seurat","",x)
  x <- gsub(" > norm.sctransform"," > sctransform",x)
  x <- gsub("coding","coding-only",x)
  x <- gsub("ribo","-ribo",x)
  gsub("Mt","-Mt",x)
}
draw(renameHrows(
  evalHeatmap(res, "clustering", what="MI", agg.by=c("classes","norm"), 
              row_split=norm, reorder_rows=TRUE, show_column_names=TRUE,
              anno_legend=FALSE, title="Mutual information") +
    evalHeatmap(res, "clustering", what="ARI", agg.by=c("classes","norm"), 
                filter=n_clus==true.nbClusts, title="ARI at true K",
                show_column_names=TRUE, anno_legend=FALSE),
  rnfn), annotation_legend_list=pipeComp:::.scaledLegend(), merge_legend=TRUE, heatmap_legend_side="bottom")
    ```

@@

Impact of restricting the type of features used on the Mutual Information (MI, left) and Adjusted Rand Index (ARI, right) of the clustering. `all` indicates that all features were used, `-Mt` stands for the exclusion of mitochondrial genes, `-ribo` the exclusion of ribosomal genes, and `coding-only` a restriction to protein-coding genes. The features were filtered out prior to normalization.

### 2020-05-12 supplementary_figures/supp3_normalization.Rmd
updated supp

```{r covar, fig.width=8, fig.height=4.5}
res <- readRDS("../data/results_norm.rds")
renameNorm(evalHeatmap(res, step="dimreduction", c("log10_total_counts","log10_total_features"), what2="meanAbsCorr.covariate2", agg.by = "norm", anno_legend=FALSE, row_split=classifyNorms(norm)))
```

>>>>

```{r covar, fig.width=8, fig.height=4.5}
res <- readRDS("../data/results_norm.rds")
renameNorm(evalHeatmap(res, step="dimreduction", c("log10_total_counts","log10_total_features"), what2="meanAbsCorr.covariate2", agg.by="norm", anno_legend=FALSE, scale=FALSE, row_split=classifyNorms(norm)))
```
